---
title: Associations Across Trial
author: Curt, 3/08/2017
output:
  html_document:
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(dplyr)
require(lme4)
require(lmerTest)
require(ggplot2)
setwd("C:/Users/cdvrmd/Box Sync/Bruce Projects/Sequential Processing/PointByPoint Processing/Created by R_ArtRej")
TrialElectLev_dat = read.delim("RT_N2_SlowWave_trial&ElecLev_R.txt")
MeanCent_dat = read.delim("RT_N2_SlowWave_MeanCent_R.txt" )
```

A datapoint will be included if:  
1) Current trial is not rejected due to artifacts.  
2) Current trial is not the first trial of the block (i.e. trial 1, 101, 201, 301 etc. are not included).  
3) Both current trial and previous trial are correct.
4) Previous trial has not been rejecred due to artifacts.

SlowWavePrev and N2Curr difference scores across Trial.
```{r data, echo = FALSE}
#**************************************************************************************************************
#  ************************Difference Scores Across Trial************************;
#SlowWavePrev and N2Curr difference scores across Trial.
N2_SW_diff_Out = lmer(N2_SW_diff ~ Trial*Prev*Curr +  
                                     (Curr*Prev|Subject)  + (1|Electrode:Subject), data = TrialElectLev_dat)
summary(N2_SW_diff_Out)

```


#RT and SWPrev difference scores across Trial.
```{r data, echo = FALSE}
#**************************************************************************************************************
#  ************************Difference Scores Across Trial************************;
RT_SW_diff_Out = lmer(RT_SW_diff ~ Trial*Prev*Curr + 
                                     (Curr*Prev|Subject) + (1|Electrode:Subject), data = TrialElectLev_dat)
summary(RT_SW_diff_Out)

```


#RT and N2Curr difference scores across Trial.
```{r data, echo = FALSE}
#**************************************************************************************************************
#  ************************Difference Scores Across Trial************************;
RT_N2_diff__Out = lmer(RT_N2_diff~ Trial*Prev*Curr +  
                                     (Curr*Prev|Subject)  + (1|Electrode:Subject), data = TrialElectLev_dat)
summary(RT_N2_diff__Out)
```

