---
title: Disaggregate analyses
author: Curt, 3/08/2017
output:
  html_document:
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(dplyr)
require(lme4)
require(lmerTest)
require(ggplot2)
setwd("C:/Users/Curt/Box Sync/Bruce Projects/Sequential Processing/PointByPoint Processing/Created by R_ArtRej")
TrialElectLev_dat = read.delim("RT_N2_SlowWave_trial&ElecLev_R.txt")
MeanCent_dat = read.delim("RT_N2_SlowWave_MeanCent_R.txt" )
```
The subject mean and subject mean centered variables have been computed for SlowWavePrev.

A datapoint will be included if:  
1) Current trial is not rejected due to artifacts.  
2) Current trial is not the first trial of the block (i.e. trial 1, 101, 201, 301 etc. are not included).  
3) Both current trial and previous trial are correct.
4) Previous trial has not been rejecred due to artifacts.

SWPrevMean indicates the between-subject effect.
SWPrevCent indicates the within-subject effect.
The random effect "SWPrevCent" allows for tests of whether the association between SWPrev and N2 and RT across trials differs across subjects at levels greater than expected by chance.

SWPrev Predict N2Curr
```{r data, echo = FALSE}
#**************************************************************************************************************
#  ************************Disaggregation************************;
#N2.
Dis_N2_Out = lmer(N2Curr ~ SWPrevMean*Prev*Curr + SWPrevCent*Prev*Curr +  
                    (SWPrevCent|Subject), data = MeanCent_dat)
summary(Dis_N2_Out)

```

SWPrev Predict RTCurr
```{r data, echo = FALSE}
#**************************************************************************************************************
#  ************************Disaggregation************************;
#RT.
Dis_RT_Out = lmer(RT ~ SWPrevMean*Prev*Curr + SWPrevCent*Prev*Curr +  
                    (SWPrevCent|Subject), data = MeanCent_dat)
summary(Dis_RT_Out)

```

