---
title: N2analyses
author: Hannah, 2/8/17
output:
  html_document:
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=140)
require(dplyr)
require(lme4)
require(lmerTest)
require(ggplot2)
```

**Major changes: Added analysis of grand averages, as well as simple slopes for change over course of experiment. Also added plots of marginal means for Current x Previous at beginning, middle, and end of experiment.**

Curt reprocessed the data using an automatic artifact rejection process with +- 75 uV as the threshold for rejection. He quantified the N2 for each trial between [put N2 interval here]. I selected the trials that fit these criteria:

“Current” trial will be included if:  
1) Current trial is not rejected due to artifacts.  
2) Current trial is not the first trial of the block (i.e. trial 1, 101, 201, 301 etc. are not included).  
3) Both current trial and previous trial are correct.  

The following electrodes are included: F3, Fz, F4, FC3, FCz, FC4, C3, Cz, C4.

The following subjects were excluded:  
- 92 (wrong number of trials, possibly data collection error)  
- 64 (<50% accuracy rate)  
- 65 (only 187 usable trials (23.3%))

One or more electrodes were bad for subjects 19, 74, and 137 (data from those electrodes is discarded).  

Besides that, ERP data is completely missing for subs 36, 37, 39, 40, 42, 43, 46, 51, 53, 54, 56, 57, 78.
  
**Total sample is 36 subjects.**

### 1. Grand averages, not looking at change over experiment

[Put grand average waveforms here]
``` {r plotGrand, echo=FALSE}
# Put ggplotting code for grand averages here


```
  
*Current and Previous are effect coded.* Uses trial-level data to test (rather than individual averages). Doesn't include Trial in model. 

``` {r test, echo=FALSE}
# to run model
N2.correct = read.delim("./AllSubs_TBTaverages_N2_Correct_withPrevious.txt")

# add effect codes for categorical variables
N2.correct$Previous = NA
N2.correct$Previous[N2.correct$prevCond == "compat"] = -1
N2.correct$Previous[N2.correct$prevCond == "incompat"] = 1

N2.correct$Current = NA
N2.correct$Current[N2.correct$Condition == "Comp"] = -1
N2.correct$Current[N2.correct$Condition == "InComp"] = 1

grand1 = lmer(MeanCurr ~ Current*Previous + 
       (Current*Previous|Subject) + (1|Electrode:Subject), data = N2.correct)

round(summary(grand1)$coefficients, digits = 3)

```


### 3. Looking at change over course of task  
**This includes all data points, *not* averaged across subject or electrode.** 

```{r plot, echo=FALSE, warning = FALSE}

ggplot(N2.correct[!(is.na(N2.correct$MeanCurr)),], aes(Trial, MeanCurr, alpha = TrialCondition, color = TrialCondition, shape = TrialCondition)) +
  geom_point() +
  geom_smooth(method = "lm", se=F, lwd = 1.3) +
  labs(x = "Trial", y = "N2 amplitude") +
  scale_shape_manual(values=c(1,1,1,1)) +
  scale_alpha_manual(values=c(.1,.1,.1,.1)) +
  scale_color_manual(values=c("blue", "dodgerblue", "darkred", "red")) +
  theme_bw() +
  coord_cartesian(ylim=c(-2.5, 5)) +
  ggtitle("N2 amplitude")
```
  
##### Simple slopes  
  
Trial is scaled to range from 0 to 8 (instead of 1 to 800) so that the betas associated with trial are a little bigger (but significance testing is unaffected by linear scaling, so the test statistics and p values will be the same as if we used the unscaled Trial variable).   

``` {r simple, echo = FALSE}
# rescale trial
N2.correct$Trial.begin = (N2.correct$Trial-2)/100
# shift trial to look at fixed effects at middle and end of task as well
N2.correct$Trial.middle = N2.correct$Trial.begin - 4
N2.correct$Trial.end = N2.correct$Trial.begin - 8

# calculate simple slopes
fm2 <- lmer(MeanCurr ~ Condition*prevCond*Trial.begin + # need to use dummy coded variables, not effect coded
       (Condition*prevCond|Subject) + (1|Electrode:Subject), data = N2.correct)

parms <- fixef(fm2) # fixed parameters from model
vcmat <- vcov(fm2) # variance/covariance matrix of fixed effects

## each row of this matrix defines entries of parms that we want to combine:
Tmat <- matrix(NA,4,8)
Tmat[1,] <- c(rep(0,3),1,0,0,0,0) # weights for estimates for current = 0, previous = 0 (current compat-previous compat)
Tmat[2,] <- c(rep(0,3),1,0,0,1,0) # weights for estimates for current = 0, previous = 1 (current compat-previous incompat)
Tmat[3,] <- c(rep(0,3),1,0,1,0,0) # weights for estimates for current = 1, previous = 0 (current incompat-previous compat)
Tmat[4,] <- c(rep(0,3),1,0,1,1,1) # weights for estimates for current = 1, previous = 1 (current incompat-previous incompat)

parest <- Tmat %*% parms          # see above for notation
newvc <- Tmat %*% vcmat %*% t(Tmat)
ses <- sqrt(diag(newvc)) 

## final results
fin = cbind(parest, ses) %>% as.data.frame()
names(fin) = c("est", "ses")

# to calculate 95% CI intervals, lower bound = m - 2*SE, upper bound = m + 2*SE

fin$lbnd = fin$est - 2*fin$ses
fin$ubnd = fin$est + 2*fin$ses

fin = format(fin, digits = 3)

# relabel for understanding
fin$Current = c("Compat", "Compat", "Incompat", "Incompat")
fin$Previous = c("Compat", "Incompat", "Compat", "Incompat")
fin$Color = c("dark blue", "dark red", "light blue", "light red")

fin = rename(fin, Estimate = est, SE = ses, ci95_lower = lbnd, ci95_upper = ubnd)

# display
fin
```  

  
#### Beginning of experiment 
The intercept and slopes of current and previous trial condition are allowed to vary by subject. The intercept is allowed to vary by electrode nested within subject.   
  
Trial is scaled to range from 0 to 8.  

Categorical variables are effect coded. 

``` {r model1, echo = FALSE}

m1 = lmer(MeanCurr ~ Current*Previous*Trial.begin + 
       (Current*Previous|Subject) + (1|Electrode:Subject), data = N2.correct)
```

**Random effects:**  
``` {r model1random, echo = FALSE}
summary(m1)$varcor
```

**Fixed effects:**
``` {r model1fixed, echo = FALSE} 
round(summary(m1)$coefficients[c(1:3,5),1:5], digits = 3)
round(summary(m1)$coefficients[c(4,6:8),1:5], digits = 3)
```

**Marginal means:**  
Error bars represent standard error.  
``` {r model1marginal, echo = FALSE}
# use dummy coded model
# was calculated earlier for simple slopes (fm2)
parms <- fixef(fm2) # fixed parameters from model
vcmat <- vcov(fm2) # variance/covariance matrix of fixed effects

## each row of this matrix defines entries of parms that we want to combine:
Tmat <- matrix(NA,4,8)
Tmat[1,] <- c(1,0,0,0,0,rep(0,3)) # weights for estimates for current = 0, previous = 0 (current compat-previous compat)
Tmat[2,] <- c(1,0,1,0,0,rep(0,3)) # weights for estimates for current = 0, previous = 1 (current compat-previous incompat)
Tmat[3,] <- c(1,1,0,0,0,rep(0,3)) # weights for estimates for current = 1, previous = 0 (current incompat-previous compat)
Tmat[4,] <- c(1,1,0,0,1,rep(0,3)) # weights for estimates for current = 1, previous = 1 (current incompat-previous incompat)

parest <- Tmat %*% parms          # see above for notation
newvc <- Tmat %*% vcmat %*% t(Tmat)
ses <- sqrt(diag(newvc)) 

## calculate estimates, standard error
fin = cbind(parest, ses) %>% as.data.frame()
names(fin) = c("est", "ses")

# to calculate 95% CI intervals, lower bound = m - 2*SE, upper bound = m + 2*SE
fin$lbnd = fin$est - 2*fin$ses
fin$ubnd = fin$est + 2*fin$ses

# format for understanding
fin$Current = c("Compat", "Compat", "Incompat", "Incompat")
fin$Previous = c("Compat", "Incompat", "Compat", "Incompat")
fin$Previous = factor(fin$Previous)
fin$Current = factor(fin$Current)
fin = rename(fin, Estimate = est, SE = ses) 

# for plotting
limits <- aes(ymax = Estimate + SE, ymin=Estimate - SE)
ggplot(fin, aes(x = Previous, y = Estimate, group = Current, color = Current)) +
  geom_line() +
  geom_point(size = 4) +
  geom_errorbar(limits, width=0.2) +
  scale_color_manual(values=c("purple", "forestgreen")) +
  ylab("N2 amplitude") +
  coord_cartesian(ylim = c(1.2, 3.4))
  
```


#### Middle of experiment 
Trial is shifted so that trial ranges from -4 to 4.  
*Just the fixed effects of Current and Previous are shown, the random effects and fixed effects of any Trial interactions will be the same as the first model.*  

**Fixed effects:**
``` {r model2, echo = FALSE}
# effect
m2 = lmer(MeanCurr ~ Current*Previous*Trial.middle + 
       (Current*Previous|Subject) + (1|Electrode:Subject), data = N2.correct)

round(summary(m2)$coefficients[c(1:3,5),1:5], digits = 3)

```
**Marginal means:**  
Error bars represent standard error.  
``` {r model2marginal, echo = FALSE}
# use dummy coded model
fm3 <- lmer(MeanCurr ~ Condition*prevCond*Trial.middle + # need to use dummy coded variables, not effect coded
       (Condition*prevCond|Subject) + (1|Electrode:Subject), data = N2.correct)
parms <- fixef(fm3) # fixed parameters from model
vcmat <- vcov(fm3) # variance/covariance matrix of fixed effects

## each row of this matrix defines entries of parms that we want to combine:
Tmat <- matrix(NA,4,8)
Tmat[1,] <- c(1,0,0,0,0,rep(0,3)) # weights for estimates for current = 0, previous = 0 (current compat-previous compat)
Tmat[2,] <- c(1,0,1,0,0,rep(0,3)) # weights for estimates for current = 0, previous = 1 (current compat-previous incompat)
Tmat[3,] <- c(1,1,0,0,0,rep(0,3)) # weights for estimates for current = 1, previous = 0 (current incompat-previous compat)
Tmat[4,] <- c(1,1,0,0,1,rep(0,3)) # weights for estimates for current = 1, previous = 1 (current incompat-previous incompat)

parest <- Tmat %*% parms          # see above for notation
newvc <- Tmat %*% vcmat %*% t(Tmat)
ses <- sqrt(diag(newvc)) 

## calculate estimates, standard error
fin = cbind(parest, ses) %>% as.data.frame()
names(fin) = c("est", "ses")

# to calculate 95% CI intervals, lower bound = m - 2*SE, upper bound = m + 2*SE
fin$lbnd = fin$est - 2*fin$ses
fin$ubnd = fin$est + 2*fin$ses

# format for understanding
fin$Current = c("Compat", "Compat", "Incompat", "Incompat")
fin$Previous = c("Compat", "Incompat", "Compat", "Incompat")
fin$Previous = factor(fin$Previous)
fin$Current = factor(fin$Current)
fin = rename(fin, Estimate = est, SE = ses) 

# for plot
limits <- aes(ymax = Estimate + SE, ymin=Estimate - SE)
ggplot(fin, aes(x = Previous, y = Estimate, group = Current, color = Current)) +
  geom_line() +
  geom_point(size = 4) +
  geom_errorbar(limits, width=0.2) +
  scale_color_manual(values=c("purple", "forestgreen")) +
  ylab("N2 amplitude") +
  coord_cartesian(ylim = c(1.2, 3.4))
```


#### End of experiment 
Trial is shifted so that trial ranges from -8 to 0.  
*Just the fixed effects of Current and Previous are shown, the random effects and fixed effects of any Trial interactions will be the same as the first model.*
  
**Fixed effects:**
``` {r model3, echo = FALSE}
m3 = lmer(MeanCurr ~ Current*Previous*Trial.end + 
       (Current*Previous|Subject) + (1|Electrode:Subject), data = N2.correct)

round(summary(m3)$coefficients[c(1:3,5),1:5], digits = 3)
```
**Marginal means:**  
Error bars represent standard error.  
``` {r model3marginal, echo = FALSE}
# use dummy coded model
fm4 <- lmer(MeanCurr ~ Condition*prevCond*Trial.end + # need to use dummy coded variables, not effect coded
       (Condition*prevCond|Subject) + (1|Electrode:Subject), data = N2.correct)
parms <- fixef(fm4) # fixed parameters from model
vcmat <- vcov(fm4) # variance/covariance matrix of fixed effects

## each row of this matrix defines entries of parms that we want to combine:
Tmat <- matrix(NA,4,8)
Tmat[1,] <- c(1,0,0,0,0,rep(0,3)) # weights for estimates for current = 0, previous = 0 (current compat-previous compat)
Tmat[2,] <- c(1,0,1,0,0,rep(0,3)) # weights for estimates for current = 0, previous = 1 (current compat-previous incompat)
Tmat[3,] <- c(1,1,0,0,0,rep(0,3)) # weights for estimates for current = 1, previous = 0 (current incompat-previous compat)
Tmat[4,] <- c(1,1,0,0,1,rep(0,3)) # weights for estimates for current = 1, previous = 1 (current incompat-previous incompat)

parest <- Tmat %*% parms          # see above for notation
newvc <- Tmat %*% vcmat %*% t(Tmat)
ses <- sqrt(diag(newvc)) 

## calculate estimates, standard error
fin = cbind(parest, ses) %>% as.data.frame()
names(fin) = c("est", "ses")

# to calculate 95% CI intervals, lower bound = m - 2*SE, upper bound = m + 2*SE
fin$lbnd = fin$est - 2*fin$ses
fin$ubnd = fin$est + 2*fin$ses

# format for understanding
fin$Current = c("Compat", "Compat", "Incompat", "Incompat")
fin$Previous = c("Compat", "Incompat", "Compat", "Incompat")
fin$Previous = factor(fin$Previous)
fin$Current = factor(fin$Current)
fin = rename(fin, Estimate = est, SE = ses) 

# for plotting
limits <- aes(ymax = Estimate + SE, ymin=Estimate - SE)
ggplot(fin, aes(x = Previous, y = Estimate, group = Current, color = Current)) +
  geom_line() +
  geom_point(size = 4) +
  geom_errorbar(limits, width=0.2) +
  scale_color_manual(values=c("purple", "forestgreen")) +
  ylab("N2 amplitude") +
  coord_cartesian(ylim = c(1.2, 3.4))
```

