#install.packages("evaluate")
#install.packages("htmltools")
#install.packages("knitr")
#install.packages("jsonlite")
#install.packages("base64enc")
#install.packages("rprojroot")
---
title: SlowWaveanalyses
author: Curt, 2/3/2017
output:
  html_document:
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=140)
require(dplyr)
require(lme4)
require(lmerTest)
require(ggplot2)
```


**Major changes: Using SlowWave data that was reprocessed to ensure correct artifact rejection processes. Appropriate electrodes were rejected for Subs 19, 74, 137 before artifact rejection process.**

I quantified the SlowWave for each trial and sent it to me as a data file. I selected the trials that fit these criteria:

“Current” trial will be included if:  
1) Current trial is not rejected due to artifacts.  
2) Current trial is not the first trial of the block (i.e. trial 1, 101, 201, 301 etc. are not included).  
3) Both current trial and previous trial are correct.  

The following electrodes are included: F3, Fz, F4, FC3, FCz, FC4, C3, Cz, C4

The following subjects were excluded:  
- 92 (wrong number of trials, possibly data collection error)  
- 64 (<50% accuracy rate)  
- 65 (only 187 usable trials (23.3%))

Besides that, SlowWave data is missing for subs 36, 37, 39, 40, 42, 43, 46, 51, 53, 54, 56, 57, 78.
  
**Total sample is 36 subjects.**

First, here's the plot showing SlowWave amplitude across trial with separate regression lines fitted for each condition. **This includes all data points, *not* averaged across subject or electrode.** 


setwd("C:/Users/psycworks/Desktop/Box Sync/Bruce Projects/Sequential Processing/PointByPoint Processing/Created by R_ArtRej/SlowWave")
path = "C:/Users/psycworks/Desktop/Box Sync/Bruce Projects/Sequential Processing/PointByPoint Processing/"


```{r plot, echo=FALSE, warning = FALSE}
setwd("C:/Users/psycworks/Desktop/Box Sync/Bruce Projects/Sequential Processing/PointByPoint Processing/Created by R_ArtRej/SlowWave")
path = "C:/Users/psycworks/Desktop/Box Sync/Bruce Projects/Sequential Processing/PointByPoint Processing/"
SlowWave.correct = read.delim("SlowWave_TBT_Cond_Wprev_3subs.txt")

ggplot(SlowWave.correct[!(is.na(SlowWave.correct$MeanCurr)),], aes(Trial, MeanCurr, alpha = TrialCondition, color = TrialCondition, shape = TrialCondition)) +
  geom_point() +
  geom_smooth(method = "lm", se=F, lwd = 1.3) +
  labs(x = "Trial", y = "SlowWave amplitude") +
  scale_shape_manual(values=c(1,1,1,1)) +
  scale_alpha_manual(values=c(.1,.1,.1,.1)) +
  scale_color_manual(values=c("blue", "dodgerblue", "darkred", "red")) +
  theme_bw() +
  coord_cartesian(ylim=c(-5, 1)) +
  ggtitle("SlowWave amplitude")
```

  
Here's the output to the MLM with current and previous trial compatibility and trial as predictors. Trial is scaled to range from 0 to 8 (instead of 1 to 800) so that the betas associated with trial are a little bigger (but significance testing is unaffected by linear scaling, so the test statistics and p values will be the same as if we used the unscaled Trial variable).   

The intercept and slopes of current and previous trial condition are allowed to vary by subject. The intercept is allowed to vary by electrode nested within subject.   

Categorical variables are effect coded. 

```{r rescaleTrial, echo=FALSE}

# rescale trial
SlowWave.correct$Trial.begin = (SlowWave.correct$Trial-2)/100
# shift trial to look at fixed effects at middle and end of task as well
SlowWave.correct$Trial.middle = SlowWave.correct$Trial.begin - 4
SlowWave.correct$Trial.end = SlowWave.correct$Trial.begin - 8

# add effect codes for categorical variables
SlowWave.correct$Previous = NA
SlowWave.correct$Previous[SlowWave.correct$prevCond == "compat"] = -1
SlowWave.correct$Previous[SlowWave.correct$prevCond == "incompat"] = 1

SlowWave.correct$Current = NA
SlowWave.correct$Current[SlowWave.correct$Condition == "Comp"] = -1
SlowWave.correct$Current[SlowWave.correct$Condition == "InComp"] = 1

```
## SlowWave results for beginning of experiment
``` {r model1, echo = FALSE}

m1 = lmer(MeanCurr ~ Current*Previous*Trial.begin + 
       (Current*Previous|Subject) + (1|Electrode:Subject), data = SlowWave.correct)

summary(m1)
```


## SlowWave results for middle of experiment  
Trial is shifted to look at fixed effects at the middle and the end of the experiment (i.e. shift so that the mean of trial is 0 and ranges from -4 to 4 and then shift so that trial ranges from -8 to 0).  
*Just the fixed effects are shown, the random effects will be the same as the first model.*

``` {r model2, echo = FALSE}
# effect
m2 = lmer(MeanCurr ~ Current*Previous*Trial.middle + 
       (Current*Previous|Subject) + (1|Electrode:Subject), data = SlowWave.correct)

round(summary(m2)$coefficients, digits = 4)

```


## SlowWave results for end of experiment
``` {r model3, echo = FALSE}
m3 = lmer(MeanCurr ~ Current*Previous*Trial.end + 
       (Current*Previous|Subject) + (1|Electrode:Subject), data = SlowWave.correct)

round(summary(m3)$coefficients, digits = 4)

```
