---
title: RT analyses
author: Hannah, 1/24/2017
output:
  html_document:
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=140)
require(dplyr)
require(ggplot2)
require(lme4)
require(lmerTest)
```

Uses RT data from Alc Caf for condition 4 (Placebo-placebo) and condition 5 (Control). Subject 92 (less than 800 trials, possible data collection errors) and 64 (less than 50% accuracy) are excluded. The sample includes 50 subjects.

“Current” trial will be included if:  
1) Current trial is not the first trial of the block (i.e. trial 1, 101, 201, 301 etc. are not included).  
2) Both current trial and previous trial are correct.  

Each subject did 800 trials, with 100 trials in each of 8 blocks. RTs were dropped if less than 100 ms and longer than 2000 ms.


```{r data, echo = FALSE}
dat = read.delim("FlankerRT_forR.txt")
dat = dat[dat$ACC == 1 & dat$PrevTrialACC == 1,]

# discard trials for which RT is less than 100 and greater than 2000
dat$RTdrop = dat$RT_Raw
dat$RTdrop[dat$RT_Raw < 100] = NA
dat$RTdrop[dat$RT_Raw > 2000] = NA

# add trial information
dat$prevTrialType = NA
dat$prevTrialType[dat$PrevTrialType == 1] = "compatible"
dat$prevTrialType[dat$PrevTrialType == 2] = "incompatible"
dat$prevTrialType = factor(dat$prevTrialType)

# select subjects in Condition 4 and 5, take out 92 and 64
dat.use = dat[dat$Condition == 4|dat$Condition == 5,] %>% 
  filter(Subject != 92) %>% 
  filter(Subject != 64)

```

## 1. Gratton effect at experiment level

```{r plotGratton, echo=FALSE}
dat.use[!(is.na(dat.use$prevTrialType)),] %>% 
  ggplot(aes(prevTrialType, RT_Winsor2SD, color = TrialType)) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(aes(group=TrialType), fun.y = mean, geom="line") +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = .1) +
  #facet_wrap(~Task, labeller=labeller(Task = facet_labels)) + 
  #  ggtitle("Total number of errors") +
  labs(y = "RT (extreme dropped)", x = "Previous trial") +
  scale_color_manual(values = c(1, 2), guide = guide_legend(title = "Current trial")) +
  coord_cartesian(ylim=c(350,450)) +
  theme_bw() +
  ggtitle("Condition 4 and condition 5")

```

Current trial type and previous trial type are effect coded.

``` {r testGratton, echo=FALSE}
# make effect coding variables
dat.use$Current.e = NA
dat.use$Current.e[dat.use$TrialType == "compatible"] = -1
dat.use$Current.e[dat.use$TrialType == "Incompatible"] = 1

dat.use$Previous.e = NA
dat.use$Previous.e[dat.use$prevTrialType == "compatible"] = -1
dat.use$Previous.e[dat.use$prevTrialType == "incompatible"] = 1


# model determined through model specification procedures (see "1 Find CAE.R)
lmer(RTdrop ~ Current.e*Previous.e + (Current.e|Subject), data = dat.use) %>% 
  summary() # use this
```

Significant interaction between current trial and previous trial indicates the presence of the Gratton effect. **Need pairwise comparisons?**

## 2. Binning analyses

Here I separated trials into four bins of 200 trials each (to match binning analysis of N2). 

``` {r binplot, echo = FALSE}
dat.use$Bin[dat.use$Block == 1|dat.use$Block == 2] = 1
dat.use$Bin[dat.use$Block == 3|dat.use$Block == 4] = 2
dat.use$Bin[dat.use$Block == 5|dat.use$Block == 6] = 3
dat.use$Bin[dat.use$Block == 7|dat.use$Block == 8] = 4

dat.use[!(is.na(dat.use$prevTrialType)),] %>% 
  ggplot(aes(prevTrialType, RT_Winsor2SD, color = TrialType)) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(aes(group=TrialType), fun.y = mean, geom="line") +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = .1) +
  facet_wrap(~Bin) + 
  #  ggtitle("Total number of errors") +
  labs(y = "RT (extreme dropped)", x = "Previous trial") +
  scale_color_manual(values = c(1, 2), guide = guide_legend(title = "Current trial")) +
  coord_cartesian(ylim=c(350,450)) +
  theme_bw() 

```

#### Bin 1 
Just shows fixed effects
``` {r bin1, echo = FALSE}
# same random effects structure as experiment model
bin1 = dat.use[dat.use$Bin == 1,] %>% 
  lmer(RTdrop ~ Current.e*Previous.e + (Current.e|Subject), .)
round(summary(bin1)$coefficients, digits = 4)
```
#### Bin 2 
Just shows fixed effects
``` {r bin2, echo = FALSE}
# same random effects structure as experiment model
bin2 = dat.use[dat.use$Bin == 2,] %>% 
  lmer(RTdrop ~ Current.e*Previous.e + (Current.e|Subject), .)
round(summary(bin2)$coefficients, digits = 4)
```
#### Bin 3 
Just shows fixed effects
``` {r bin3, echo = FALSE}
# same random effects structure as experiment model
bin3 = dat.use[dat.use$Bin == 3,] %>% 
  lmer(RTdrop ~ Current.e*Previous.e + (Current.e|Subject), .)
round(summary(bin3)$coefficients, digits = 4)
```
#### Bin 4 
Just shows fixed effects
``` {r bin4, echo = FALSE}
# same random effects structure as experiment model
bin4 = dat.use[dat.use$Bin == 4,] %>% 
  lmer(RTdrop ~ Current.e*Previous.e + (Current.e|Subject), .)
round(summary(bin4)$coefficients, digits = 4)
```

## 3. Trial level analyses across course of task
Here I'm using all accepted trials (see criteria above) for all subjects across the course of the experiment (across 800 trials). Plotted data points are NOT summarized over subjects.

``` {r trialLevel, echo = FALSE, warning = FALSE}
# data set up in 8 blocks, trial goes 1-100 in each block
# add column for trial number across whole experiment so it goes 1-800
dat.use$ExpTrial = dat.use$Trial + (100*(dat.use$Block - 1))

# add label for trial condition
dat.use$TrialCondition = "Previous compatible - Current incompatible"
dat.use$TrialCondition[dat.use$prevTrialType == "compatible" & dat.use$TrialType == "compatible"] = 
  "Previous compatible - Current compatible"
dat.use$TrialCondition[dat.use$prevTrialType == "incompatible" & dat.use$TrialType == "Incompatible"] = 
  "Previous incompatible - Current incompatible"
dat.use$TrialCondition[dat.use$prevTrialType == "incompatible" & dat.use$TrialType == "compatible"] = 
  "Previous incompatible - Current compatible"
dat.use$TrialCondition = factor(dat.use$TrialCondition)

# summarise across subjects for plotting
ggplot(dat.use, aes(ExpTrial, RTdrop, alpha = TrialCondition, color = TrialCondition, shape = TrialCondition)) +
  geom_point() +
  geom_smooth(method = "lm", se=F, lwd = 1.3) +
  labs(x = "Trial", y = "RT (dropped if <100 and >2000)") +
  scale_shape_manual(values=c(1,19,1,19)) +
  scale_alpha_manual(values=c(.1,.1,.1,.1)) +
  scale_color_manual(values=c("blue", "dodgerblue", "darkred", "red")) +
  theme_bw() +
  coord_cartesian(ylim=c(325, 500)) 
```

##### Simple slopes
``` {r simple, echo = FALSE}
fm2 <- lmer(RTdrop ~ TrialType*prevTrialType*ExpTrial + (TrialType|Subject), data = dat.use)

parms <- fixef(fm2) # fixed parameters from model
vcmat <- vcov(fm2) # variance/covariance matrix of fixed effects

## each row of this matrix defines entries of parms that we want to combine:
Tmat <- matrix(NA,4,8)
Tmat[1,] <- c(rep(0,3),1,0,0,0,0) # weights for estimates for current = 0, previous = 0 (current compat-previous compat)
Tmat[2,] <- c(rep(0,3),1,0,0,1,0) # weights for estimates for current = 0, previous = 1 (current compat-previous incompat)
Tmat[3,] <- c(rep(0,3),1,0,1,0,0) # weights for estimates for current = 1, previous = 0 (current incompat-previous compat)
Tmat[4,] <- c(rep(0,3),1,0,1,1,1) # weights for estimates for current = 1, previous = 1 (current incompat-previous incompat)

parest <- Tmat %*% parms          # see above for notation
newvc <- Tmat %*% vcmat %*% t(Tmat)
ses <- sqrt(diag(newvc)) 

## final results
fin = cbind(parest, ses) %>% as.data.frame()
names(fin) = c("est", "ses")

# to calculate 95% CI intervals, lower bound = m - 2*SE, upper bound = m + 2*SE

fin$lbnd = fin$est - 2*fin$ses
fin$ubnd = fin$est + 2*fin$ses

fin = format(fin, digits = 3)

# relabel for understanding
fin$Previous = c("Compat", "Compat", "Incompat", "Incompat")
fin$Current = c("Compat", "Incompat", "Compat", "Incompat")
fin$Color = c("dark blue", "light blue", "dark red", "light red")

fin = rename(fin, Estimate = est, SE = ses, ci95_lower = lbnd, ci95_upper = ubnd)

# display
fin
```  

#### Beginning of experiment
Trial ranges from 0-798 to estimate effects of previous and current trial at the beginning of the experiment. Categorical variables are effect coded.
``` {r begin, echo = FALSE}
dat.use$Tr.begin = dat.use$ExpTrial - 2

begin = lmer(RTdrop ~ Current.e*Previous.e*Tr.begin + (Current.e|Subject), data = dat.use)

round(summary(begin)$coefficients, digits = 4)
```
  
##### Summary
**At the beginning of the experiment:**  
1. The effect of current trial is significant (compatible is faster than incompatible)  
2. No effect of previous trial  
3. No interaction  

**Over the course of the experiment:**  
1. Significant effect of trial (RT getting faster over time)  
2. Significant trial x Current trial interaction (effect of current trial decreases over time)  


#### Middle of experiment
Trial ranges from -399 to 399 to estimate effects of previous and current trial at the middle of the experiment. Categorical variables are effect coded.
  
*Effect of Trial, Current x Trial, Previous x Trial, and Current x Previous x Trial are still being estimated, but not being displayed. They will match the estimates from the first model exactly.*
``` {r middle, echo = FALSE}
dat.use$Tr.middle = dat.use$Tr.begin - 399

middle = lmer(RTdrop ~ Current.e*Previous.e*Tr.middle + (Current.e|Subject), data = dat.use)

round(summary(middle)$coefficients[c(1:3,5),1:5], digits = 4)
```
  
##### Summary
**At the middle of the experiment:**  
1. The effect of current trial is still significant (compatible is faster than incompatible)  
2. Still no effect of previous trial  
3. Now significant effect of current x previous (p < .002)  


#### End of experiment
Trial ranges from -798 to 0 to estimate effects of previous and current trial at the end of the experiment. Categorical variables are effect coded.

*Effect of Trial, Current x Trial, Previous x Trial, and Current x Previous x Trial are still being estimated, but not being displayed. They will match the estimates from the first model exactly.*
``` {r end, echo = FALSE}
dat.use$Tr.end = dat.use$Tr.begin - 798

end = lmer(RTdrop ~ Current.e*Previous.e*Tr.end + (Current.e|Subject), data = dat.use)

round(summary(end)$coefficients[c(1:3,5),1:5], digits = 4)
```
  
##### Summary
**At the end of the experiment:**  
1. The effect of current trial is still significant (compatible is faster than incompatible)  
2. Still no effect of previous trial  
3. Now marginally effect of current x previous (p < .097)  
